---
title: "PPE ABC Algorithm - Documentatin"
output: html_notebook
---

#Load in requirments for Sample
```{r}
#Load the required packages
library(dplyr)
library(ggplot2)
library(truncnorm)
library(data.table)   
HDIofMCMC             = function( sampleVec , credMass=0.95 ) {
  # Computes highest density interval from a sample of representative values,
  #   estimated as shortest credible interval.
  # Arguments:
  #   sampleVec
  #     is a vector of representative values from a probability distribution.
  #   credMass
  #     is a scalar between 0 and 1, indicating the mass within the credible
  #     interval that is to be estimated.
  # Value:
  #   HDIlim is a vector containing the limits of the HDI
  sortedPts = sort( sampleVec )
  ciIdxInc = ceiling( credMass * length( sortedPts ) )
  nCIs = length( sortedPts ) - ciIdxInc
  ciWidth = rep( 0 , nCIs )
  for ( i in 1:nCIs ) {
    ciWidth[ i ] = sortedPts[ i + ciIdxInc ] - sortedPts[ i ]
  }
  HDImin = sortedPts[ which.min( ciWidth ) ]
  HDImax = sortedPts[ which.min( ciWidth ) + ciIdxInc ]
  HDIlim = c( HDImin , HDImax )
  return( HDIlim )
}
source("Truncated_Beta_Distribution_Functions.R")
```
Here we load the neccessary requirements for needed for the simulation

```{r}
d1a            = read.csv( "/Users/michaelcollins/Documents/Predictive Performance Equation/AHA_Data/UDRI_Data_PPE_Variables_var2.csv" )
Subject        = d1a[d1a$Subject %in% unique(d1a$Subject)[1:5], ]
Subject        = arrange(Subject, user_id, ID)

```
Here we load in the data that will be used for the simulation

```{r}
Instances = 
  unique(
  Subject %>%
  group_by(ID ) %>%
  mutate(Marker = length(ID)) %>%
  select(ID, Marker)
 ) 

People = 
  unique(
    Subject %>%
      group_by(user_id ) %>%
      mutate(Marker = length( unique(ID) ) ) %>%
      select(user_id, Marker)
  ) 

Subject$Diff_Post = 0
Subject$tau_Post  = 0
Subject$s_Post    = 0
Subject$m_post    = 0

```
Here creates two variables "instances" and "People", which represent the number unique skills complted by each participant and how many instances are in each skill completed by a participant. This is used to provide each participant skills combination with the correct number of parameters. 

```{r}
#Set the Tuning parameter
Tuning              = .05
NumberofSubjects    = length(unique(Subject$user_id))
NumberofItems       = length(unique(Subject$ID))
NumberofSamples     = 5000
setwd("/Users/michaelcollins/Documents/Thoughts/Aproximate Bayesian Computation")
File_Name = "PPE_Results.txt"

```
Now we set the some of the sampling specific parameters for the simulation
-Tuning, controls the the accepction and refection specifictiy in the ABC sampling
-NumberofSubjects: Number of participants
-NumberofItem: number of skills that were done by participants
-set the working directory
-Give the dataoutput in the data file

#ABC Model
##Bayesian Models
###Population Distribution
```{r}
  print(i)
  #Population Level Distribution
  Population_tau = rtbeta(1 , .85 * 10  , (1-.85)  * 10,  lower = .01, upper = 1 )
  Population_s   = rtbeta(1 , .05 * 10  , (1-.05)  * 10,  lower = .01, upper = 1 )
  Population_b   = rtbeta(1 , .01 * 25 , (1-.01)   * 25,  lower = .01, upper = .25 )
  Population_m   = rtbeta(1 , .01 * 25 , (1-.01)   * 25,  lower = .01, upper = .25 )
  
  Population_tau_variation  = runif(1, 100, 1000)
  Population_s_variation    = runif(1, 100, 1000)
  Population_b_variation    = runif(1, 100, 1000)
  Population_m_variation    = runif(1, 100, 1000)

```

Here we specity the the population parameters for the model. The population parameters are made up of two different distribution. For each free parameter, there is a distribution there is a one distribution for the mean and one distribution for the certinty. 
-mean parameters are represeted as beta distributions
-certinty parameters are rpreseted as uniform distributions. 


####Metroplitis sampling
```{r}
  if(i > 1){
    Prev_tau = unique(Subject$Population_tau_mean_Post)
    Prev_tau_variation = unique(Subject$Population_tau_sd_Post)
    #------
    Prev_tau_LL = prod( dtbeta( unique(Subject$Subject_tau_mean_Post), Prev_tau * Prev_tau_variation, (1-Prev_tau) * Prev_tau_variation, lower = 0, upper = 1 ) )
         tau_LL = prod( dtbeta( unique(Subject$Subject_tau_mean_Post), Population_tau * Population_tau_variation, (1-Population_tau) * Population_tau_variation, lower = 0, upper = 1 ) )
    tau_r       = tau_LL/ Prev_tau_LL
   #---------------------------------------------------------------------------------------------------------------
    Prev_s = unique(Subject$Population_s_mean_Post)
    Prev_s_variation = unique(Subject$Population_s_sd_Post)
    #------
    Prev_s_LL = prod( dtbeta( unique(Subject$Subject_s_mean_Post), Prev_s * Prev_s_variation, (1-Prev_s) * Prev_s_variation, lower = 0, upper = 1 ) )
    s_LL = prod( dtbeta(      unique(Subject$Subject_s_mean_Post), Population_s * Population_s_variation, (1-Population_s) * Population_s_variation, lower = 0, upper = 1 ) )
    s_r       = s_LL/ Prev_s_LL
    #---------------------------------------------------------------------------------------------------------------
    Prev_b = unique(Subject$Population_b_mean_Post)
    Prev_b_variation = unique(Subject$Population_b_sd_Post)    
    #------
    Prev_b_LL = prod( dtbeta( unique(Subject$Subject_b_mean_Post), Prev_b * Prev_b_variation, (1-Prev_b) * Prev_b_variation, lower = 0, upper = 1 ) )
    b_LL = prod( dtbeta(      unique(Subject$Subject_b_mean_Post), Population_b * Population_b_variation, (1-Population_b) * Population_b_variation, lower = 0, upper = 1 ) )
    b_r       = b_LL/ Prev_b_LL
    #---------------------------------------------------------------------------------------------------------------
    Prev_m = unique(Subject$Population_m_mean_Post)
    Prev_m_variation = unique(Subject$Population_m_sd_Post)
    #------
    Prev_m_LL = prod( dtbeta( unique(Subject$Subject_m_mean_Post), Prev_m * Prev_m_variation, (1-Prev_m) * Prev_m_variation, lower = 0, upper = 1 ) )
    m_LL = prod( dtbeta(      unique(Subject$Subject_m_mean_Post), Population_m * Population_m_variation, (1-Population_m) * Population_m_variation, lower = 0, upper = 1 ) )
    m_r       = m_LL/ Prev_m_LL
    #---------------------------------------------------------------------------------------------------------------
    
  if( runif(1, 0,1) > tau_r ) {Population_tau = Prev_tau  ; Population_tau_variation = Prev_tau_variation }  
  if( runif(1, 0,1) > s_r   ) {Population_s   = Prev_s   ; Population_s_variation   = Prev_s_variation  }  
  if( runif(1, 0,1) > b_r   ) {Population_b   = Prev_b   ; Population_b_variation   = Prev_b_variation   } 
  if( runif(1, 0,1) > m_r   ) {Population_m   = Prev_m   ; Population_m_variation   = Prev_m_variation   } 
    
  }
```

After being initialized ($i  > 1$) we use a Metroplois algorithm is used to sample parameters. 
Metropolis algorithm works by comparing a parameter values sampled from the prior to the previous posterior value. The likilehood of these two values for accounting the lower values in the hierarchy. A likelihood ratio is computed  and an acceptance probability for these new parameter values. 

###Subject Level
```{r}
 Subject_tau = rtbeta(NumberofSubjects, Population_tau * Population_tau_variation  , (1 - Population_tau)  * Population_tau_variation ,  lower = .01, upper = 1 )
  Subject_s   = rtbeta(NumberofSubjects, Population_s   * Population_s_variation    , (1 - Population_s)  * Population_s_variation     ,  lower = .01, upper = 1 )
  Subject_b   = rtbeta(NumberofSubjects, Population_b   * Population_b_variation    , (1 - Population_b)  * Population_b_variation     ,  lower = .01, upper = .25 )
  Subject_m   = rtbeta(NumberofSubjects, Population_m   * Population_m_variation    , (1 - Population_m)  * Population_m_variation     ,  lower = .01, upper = .25 )
  
  Subject_tau_variation  = runif(NumberofSubjects, 100, 1000)
  Subject_s_variation    = runif(NumberofSubjects, 100, 1000)
  Subject_b_variation    = runif(NumberofSubjects, 100, 1000)
  Subject_m_variation    = runif(NumberofSubjects, 100, 1000)
  #------------------------------------------------------------------------------------
  tau_mean_list  = rep(Subject_tau         , times = People$Marker)
  tau_sd_list  = rep(Subject_tau_variation , times = People$Marker)
  #----
  s_mean_list  = rep(Subject_s            , times = People$Marker)
  s_sd_list    = rep(Subject_s_variation  , times = People$Marker)
  #----
  b_mean_list  = rep(Subject_b            , times = People$Marker)
  b_sd_list    = rep(Subject_b_variation  , times = People$Marker)
  #----
  m_mean_list  = rep(Subject_m            , times = People$Marker)
  m_sd_list    = rep(Subject_m_variation  , times = People$Marker)
  #------------------------------------------------------------------------------------
  Item_list = rep(Instances$ID, times = Instances$Marker )

```

###Item level
```{r}
 Item_tau    = rtbeta(length(tau_mean_list) , tau_mean_list * tau_sd_list , (1 - tau_mean_list) * tau_sd_list, lower = 0, upper = 1 )
  Item_s      = rtbeta(length(s_mean_list)   , s_mean_list   * s_sd_list   , (1 - s_mean_list)   * s_sd_list, lower = 0, upper = .1 )
  Item_b      = rtbeta(length(b_mean_list)   , b_mean_list   * b_sd_list   , (1 - b_mean_list)   * b_sd_list, lower = 0, upper = .25 )
  Item_m      = rtbeta(length(m_mean_list)   , m_mean_list   * m_sd_list   , (1 - m_mean_list)   * m_sd_list, lower = 0, upper = .25 )
  
  Item_tau    = rep(Item_tau , times = Instances$Marker )
  Item_s      = rep(Item_s   , times = Instances$Marker )
  Item_b      = rep(Item_b   , times = Instances$Marker )
  Item_m      = rep(Item_m   , times = Instances$Marker )
  
  Subject$tau = Item_tau
  Subject$s   = Item_s
  Subject$b   = Item_b
  Subject$m   = Item_m

```

###PPE Model
```{r}
 Subject$decay     = Subject$b + Subject$m * Subject$CumAveInvLogTime_diff_lag1
  Subject$Learning  = Subject$N^.1
  Subject$M         = Subject$Learning * Subject$T^ -Subject$decay
  Subject$Prob      = 1 / ( 1 + exp( (Subject$tau - Subject$M) /Subject$s ) )
```

#Acceptance Probability
```{r}
  #Computer Difference Measures and alpha Values
  Subject$Diff = sqrt(mean( (Subject$performance - Subject$Prob)^2 ))
  
  #Acception_Paramters$Diff  = sum(Acception_Paramters$Diff)
  p                         = runif(1, 0,1)  
  Subject$p                 = p
  Subject$alpha             = dnorm(Subject$Diff/Tuning) / dnorm(Subject$Diff_Post/Tuning)
  Accept                    = unique(Subject$p < Subject$alpha)

```

#Record the data
```{r}
   Subject$Sample       = i
    if(unique(Accept) == TRUE){
      #---------------------------------------------------------------------------------------------------
      Subject$Population_tau_mean_Post  = Population_tau
      Subject$Population_s_mean_Post    = Population_s
      Subject$Population_b_mean_Post    = Population_b
      Subject$Population_m_mean_Post    = Population_m
      #---------------------------------------------------------------------------------------------------
      Subject$Population_tau_sd_Post  = Population_tau_variation
      Subject$Population_s_sd_Post    = Population_s_variation
      Subject$Population_b_sd_Post    = Population_b_variation
      Subject$Population_m_sd_Post    = Population_m_variation
      #---------------------------------------------------------------------------------------------------
      Subject$Subject_tau_mean_Post = tau_mean_list
      Subject$Subject_s_mean_Post   =   s_mean_list
      Subject$Subject_b_mean_Post   =   b_mean_list
      Subject$Subject_m_mean_Post   =   m_mean_list
      #--------------------------------------------
      Subject$Subject_tau_sd_Post = tau_sd_list
      Subject$Subject_s_sd_Post   =   s_sd_list
      Subject$Subject_b_sd_Post   =   b_sd_list
      Subject$Subject_m_sd_Post   =   m_sd_list
      #---------------------------------------------------------------------------------------------------
      Subject$Item_tau_Post  = Item_tau
      Subject$Item_s_Post    = Item_s
      Subject$Item_b_Post    = Item_b
      Subject$Item_m_Post    = Item_m
      #---------------------------------------------------------------------------------------------------
      Subject$Perf_Post         = Subject$Prob
      Subject$Diff_Post         = Subject$Diff
    }

```
